---
title: "Transmission analysis"
author: "Chris BC"
date: "19 October 2015"
output: pdf_document
---

Analyse any relationship between transmission type and fuel efficiency as measured in miles per gallon.

Initially let's look at a simple linear regression between the two variables:
```{r}

fit1 <- lm(mpg ~ am, mtcars)
summary(fit1)

```

This suggests a strong positive relationship between the two variables. am coefficient 7.245 is well removed from 2*std. error (1.764), and t-value probability approaches zero, indicating a strong probability.

Residual standard error appears significant so let's look at residuals.

```{r}

autos<-mtcars[which(mtcars$am==0),]
manuals<-mtcars[which(mtcars$am==1),]
library(ggplot2)
plot(mtcars$am, resid(fit1))
plot(autos$mpg, resid(fit1)[which(mtcars$am==0)], col="red", xlim=c(10,35),ylim=c(-10,10))
points(manuals$mpg, resid(fit1)[which(mtcars$am==1)], col="blue")
```

It is clear that there's a strong correlation between the two variables, but with residuals increasing as mpg increases there's more to the story so a more detailed model is required.

```{r}

summary(glm(mpg~.,data=mtcars))
plot(mtcars$mpg, resid(glm(mpg~.,data=mtcars)))

```

This shows a series of residuals lacking a pattern, indicating a candidate model, but with all of our coefficients being within 2 standard errors of zero fails to provide any insight.

```{r}

fit2<-glm(mpg~am+cyl+wt,data=mtcars)
fit3<-glm(mpg~am+hp+wt,data=mtcars)
summary(fit2)
summary(fit3)
plot(mtcars$mpg, resid(fit2))
plot(mtcars$mpg,resid(fit3))

```

These models show a good spread of residuals but fails to show a correlation (within 2s.d. of 0) between mpg and transmission.

As shown below both poisson and binomial models olso present models that appear to have high certainty but which have a strong pattern in their residuals.

```{r}
fitP<-glm(mpg~am,data=mtcars,family="poisson")
fitP2<-glm(mpg~.,data=mtcars,family="poisson")
fitB<-glm(am~mpg,data=mtcars,family="binomial")
summary(fitP)
summary(fitP2)
summary(fitB)
plot(mtcars$mpg,resid(fitP), col="red")
points(mtcars$mpg,resid(fitB),col="blue")
legend(26,-1,c("Poisson model","Binomial model"),pch=c(1,1),col=c("red","blue"))

```